Variants on the Flight and Maintenance Planning problem

### 2018-07-13 latex/notes/complexite/FMP_variants.tex

    The objective of the problem discussed by \cite{Brunner2013} is to minimise the size of the workforce. Apart from coverage requirements, expressed as a minimum ($b_{t}$), there are three other constraints. The first constraint limits the number of days worked to a maximum value ($\overline{D}^{work}$). The second and third constraints define a range on the number of consecutive days worked ($\underline{D}^{on} - \overline{D}^{on}$) and days-off ($\underline{D}^{off} - \overline{D}^{off}$), respectively.

    All missions can use any aircraft (homogeneous fleet). 
    Even if we have a fixed (not a range) number of consecutive non-working days, the original proof was made with both ranges being fixed.

    Now, a simplified description of the original problem is done that matches this problem:

    For each period $t$ with demand $b_{t}$, we create a task $j$ with $T_j = \{t\}$ that requires $r_{j}$ resources.
    For each worker $i$ we have a corresponding resource $i$.
    Since the workforce has the same skills, we assume $O_i = \{j \in \mathca{J}\}$.

    A series of $j \in \mathcal{J}$ tasks are planned along a horizon divided into $t \in \mathcal{T}$ periods. Since all tasks are already scheduled, we know time periods $T_j \subset \mathcal{T}$ in which they will be realized. Similarly, all the tasks to be realized in period $t$ are known and defined by set $J_t \subset \mathcal{J}$. Each task requires a certain number $r_{j}$ of resources $i \in \mathcal{I}$ which it employs for a time duration defined by $h_j$ in each period. Set $a_{j} \subset \mathcal{I}$ lists the resources that can be assigned to each task and set $O_i \subset \mathcal{J}$ consists of tasks for which resource $i$ can be used. The number of resources in use for each period $d_t$ is also known.

    The notation follows closely the \cite{Smet2015} 2.6.1 example.

@@

    \begin{itemize}
        \item The missions consume 0 hours of aircraft.
        \item The set of $j \in \mathcal{J}$ tasks
        \item No minimum duration for mission assignments.
    \end{itemize}

@@

    \begin{tabular}{ll}
        $t \in \mathcal{T}$ & Planning horizon periods. \\
        $i \in \mathcal{I}$ & Resources. \\
    \end{tabular}

@@

    \begin{tabular}{ll}
        $\overline{D}^{work}$ & Maximum number of total working days in period. \\
        $D^{off}$ & Number of consecutive non-working days. \\
        $\overline{D}^{on}$ & Maximum number of consecutive working days. \\
        $\underline{D}^{on}$ & Minimum number of consecutive working days. \\
        $b_{t}$ & Need of resources at period $t$. \\
    \end{tabular}

    \begin{itemize}
        \item $x_{it}$: resource $i$ is working on period $t$.
        \item $y_{i}$: resource $i$ is working on at least one period.
    \end{itemize}

    \begin{align}
        & \text{Min}\; \sum_{i \in \mathcal{I}} y_{i} \label{min_num_resources}
    \end{align}

    \begin{align}
        & \sum_{i \in \mathcal{I}} x_{it} \geq b_t & t \in \mathcal{T} \label{demand3} \\
        & \sum_{t \in \mathcal{T}} x_{it} \leq \overline{D}^{work} & i \in \mathcal{I} \label{max_work} \\
        & \sum_{t' = t}^{\overline{D}^{on}+t-1} x_{it'} \leq \overline{D}^{on} & i \in \mathcal{I}, t \in \mathcal{T} \label{max_work_days} \\
        & \sum_{t' = t}^{\underline{D}^{on}+t-1} x_{it'} \geq \underline{D}^{on} (x_{it} - x_{it-1}) & i \in \mathcal{I}, t \in \mathcal{T} \label{min_work_days} \\
        & \sum_{t' = t}^{D^{off}+t-1} (1 - x_{it'}) \leq D^{off} & i \in \mathcal{I}, t \in \mathcal{T} \label{max_no_work_days} \\
        & \sum_{t' = t}^{D^{off}+t-1} (1 - x_{it'}) \geq D^{off} (- x_{it} + x_{it-1}) & i \in \mathcal{I}, t \in \mathcal{T} \label{min_no_work_days} \\
        & y_{i} \geq x_{it} & i \in \mathcal{I}, t \in \mathcal{T} \label{num_resources} \\
        & x_{it}, y_{i} \in \{0, 1\} & i \in \mathcal{I}, t \in \mathcal{T} \label{binary}
    \end{align}

    Constraints (\ref{demand3}) guarantee the necessary number of workers is available to work. 
    Constraints ((\ref{max_work}) checks that the total number of working days for each worker is less than the upper limit.
    Constraints (\ref{max_work_days} - \ref{min_work_days}) control the resources meet the range of working days. Constraints (\ref{max_no_work_days} - \ref{min_no_work_days}) control the resources meet the fixed number of non-working days.
    Finally constraints (\ref{num_resources}) quantify the number of necessary workers in the planning to be minimized in (\ref{min_num_resources}).

>>>>

    \subsection{Base problem}

    The problem is described in \cite{Smet2015}, almost verbatim, as the following:

    The objective of the problem discussed by \cite{Brunner2013} is to minimize the size of the workforce. Apart from coverage requirements, expressed as a minimum ($b_{k}$), there are three other constraints. The first constraint limits the number of days worked to a maximum value ($\overline{D}^{work}$). The second and third constraints define a range on the number of consecutive days worked ($\underline{D}^{on} - \overline{D}^{on}$) and days-off ($\underline{D}^{off} - \overline{D}^{off}$), respectively.

    \subsection{Reduction}

    Now, a simplified description of our original problem is done so that it complies with the above one. For reference, see the original description of the problem at the beginning of this document. It's an adaption of the MOSIM description.

    For each period $k$ from the original problem, we create a period $t \in \mathcal{T}$. We can use the elements in these two sets indistinctly.
    For each period $k$ with demand $b_{k}$, we create a task $j$ with $T_j = \{t\}$ that requires $r_{j} = b_{t}| t \in T_j$ resources. This implies that $|J_t| = 1$ and $d_t = b_{t}$
    For each worker $i$ we have a corresponding resource $i$.
    Since all resources can cover the demand, we assume $O_i = \{j \in \mathcal{J}\}$ and $a_{j} = \{i \in \mathcal{I}\}$.
    All resources start the planning horizon have just exited a maintenance in the previous period. This implies $n_t = 0$.
    The duration of the maintenance will be equal to $M = \overline{D}^{off} = \underline{D}^{off}$. This is assumption is valid because the original proof was done with fixed values instead of ranges.
    The maximum remaining elapsed time after a maintenance is equal to $E = \overline{D}^{off}$. The minimum elapsed time before a maintenance is equal to $e=\underline{D}^{off}$.
    The usage time for each resource by any task will be $h_j = 0$ and the remaining usage time after a maintenance is $H = 0$. The minimum number of maintenances for each resource will be given by $\lceil(|\mathcal{T}| - \overline{D}^{work}) / M \rceil$.

    The model notation follows closely the \cite{Smet2015} 2.6.1 example.

@@

@@

    \begin{tabular}{ll}
        $t \in \mathcal{T}$ & Planning horizon periods. \\
        $i \in \mathcal{I}$ & Resources. \\
        $j \in \mathcal{J}$ & tasks. 
    \end{tabular}

    \begin{tabular}{ll}
        $m$ & Minimum number of maintenances per resource in planning horizon. \\
        $M$ & Duration of maintenance. \\
        $E$ & Maximum elapsed time before assigning a maintenance. \\
        $e$ & Minimum elapsed time before assigning a maintenance. \\
        $r_{j}$ & Need of resources for task $j$. \\
    \end{tabular}

    (here we could, in theory just copy the same model than we had in MOSIM).

@@

    % \subsection{Variables}

    % \begin{itemize}
    %     \item $x_{it}$: resource $i$ is working on period $t$.
    %     \item $y_{i}$: resource $i$ is working on at least one period.
    % \end{itemize}

    % \subsection{Model}

    % \begin{align}
    %     & \text{Min}\; \sum_{i \in \mathcal{I}} y_{i} \label{min_num_resources}
    % \end{align}

    % \begin{align}
    %     & \sum_{i \in \mathcal{I}} x_{it} \geq b_t & t \in \mathcal{T} \label{demand3} \\
    %     & \sum_{t \in \mathcal{T}} x_{it} \leq \overline{D}^{work} & i \in \mathcal{I} \label{max_work} \\
    %     & \sum_{t' = t}^{\overline{D}^{on}+t-1} x_{it'} \leq \overline{D}^{on} & i \in \mathcal{I}, t \in \mathcal{T} \label{max_work_days} \\
    %     & \sum_{t' = t}^{\underline{D}^{on}+t-1} x_{it'} \geq \underline{D}^{on} (x_{it} - x_{it-1}) & i \in \mathcal{I}, t \in \mathcal{T} \label{min_work_days} \\
    %     & \sum_{t' = t}^{D^{off}+t-1} (1 - x_{it'}) \leq D^{off} & i \in \mathcal{I}, t \in \mathcal{T} \label{max_no_work_days} \\
    %     & \sum_{t' = t}^{D^{off}+t-1} (1 - x_{it'}) \geq D^{off} (- x_{it} + x_{it-1}) & i \in \mathcal{I}, t \in \mathcal{T} \label{min_no_work_days} \\
    %     & y_{i} \geq x_{it} & i \in \mathcal{I}, t \in \mathcal{T} \label{num_resources} \\
    %     & x_{it}, y_{i} \in \{0, 1\} & i \in \mathcal{I}, t \in \mathcal{T} \label{binary}
    % \end{align}

    % Constraints (\ref{demand3}) guarantee the necessary number of workers is available to work. 
    % Constraints ((\ref{max_work}) checks that the total number of working days for each worker is less than the upper limit.
    % Constraints (\ref{max_work_days} - \ref{min_work_days}) control the resources meet the range of working days. Constraints (\ref{max_no_work_days} - \ref{min_no_work_days}) control the resources meet the fixed number of non-working days.
    % Finally constraints (\ref{num_resources}) quantify the number of necessary workers in the planning to be minimized in (\ref{min_num_resources}).
